Dynamical Properties of Molecular Cloud Complexes in Galaxies at the Epoch of Reionization

### 2019-05-31 WRITEUP/ms.tex

The most massive MCC is excluded in the discussion of MCC dynamics in the remainder of this paper
since its dynamics is dominated by large-scale shear.
Considering only the MCCs identified at the highest density threshold, the mass of MCCs is $\simeq10^{6.5}$\,\Msun.
We note that similarly massive molecular structures (few times 10$^7$\,\Msun\footnote{See \S{3.1} of \citet{Behrendt16a}.}) 
have been reported in idealized closed-box isolated galaxy simulations done at higher resolution (e.g., a maximum resolution of 3\,pc in the 48\,kpc box studied by \citealt{Behrendt16a}).
However, the IGM and merger and accretion histories are not properly modeled in such simulations since they adopt non-cosmological initial conditions. That said, the similar mass range found in MCCs of \flower\ is reassuring --- our results are not far off in spite of the limited resolution ($l_{\rm cell}\simeq$\,30\,pc).

@@

If constant column density (Larson's third law) is truly a fundamental properties of molecular clouds % or the roughly linear relationship between $\rho$ and $R$ of MCCs.
(i.e., if virialization is a universal property of molecular clouds with negligible surface pressure), 
one would expect a single point in the $\sigma_{\rm gas}^2/R$--$\Sigma$ relation at a given gas mass surface density $\Sigma$. 
The variation in $\sigma_{\rm gas}^2/R$ with $\Sigma$ seen here suggests that column density is not a fundamental properties of 
molecular clouds. Previous studies suggest that the observed mass-size relation of molecular clouds 
is solely a result of $\rho\propto M R^{-3}$ and may be an artifact of the limited range of 
column densities a specific molecular line tracer is sensitive to \citep[][]{Ballesteros02a, Ballesteros11a}.
%For instance, \citet{Ballesteros02a} show that the density
%of clouds ($\rho$) is independent of their size ($R$) when one properly extract these quantities in three-dimensional space instead of
%two-dimensional space; the latter is done in \obs.
%When they project the clouds into two-dimensional space and measure the densities and sizes of clouds for a given
%line tracer, they find $\rho \propto R^{-1}$, which is unsurprising since $N_H$ is just a constant.
%They therefore argue that the mass-size relation is simply a result of $\rho\propto M R^{-3}$
%observed in 2D-space at a fixed column density (cf. an actual physical relation of $M \propto R^{4/3}$ claimed in
%observational studies). % where $\rho$ would be proportional to $R^{-1/2}$
%% since $M$ would then scale as $R^{3}$ (ånd naturally display a linear relationship in $\log M - \log R$ plot).

@@

% The pressure of MCCs identified at the lowest $n_{\rm cut}$ ranges between $\log\left(P/K$\,\cc$\right)\in[6, 8.7]$, with a median of 7.4.
The pressure of MCCs identified at the highest $n_{\rm cut}$ ranges between $\log(P/k_B)\in[7, 9]$\,K\,\cc, with a median of 7.6\,K\,\cc.
Such high pressure is comparable to those observed in local ULIRGs; however, the molecular clouds in the latter are concentrated within their central regions \citep{Downes98a, Sakamoto08a}. In our simulated galaxy, the high pressure MCCs of \flower are found throughout the disk. This difference likely stems from the different physical mechanisms giving rise to the formation (and thus the nature) of these molecular structures.
%
In local ULIRGs, MCCs are likely formed by shock compression and cloud-cloud collisions that funnel large amounts of gas from the progenitor galaxies towards the central region \citep{Tan00a, Wu18a}. On the other hand, the highly turbulent structures of \flower are the result of extra-planar flows \citep{Kohandel19a} and higher velocity accretion/SN-driven outflows \citep{Gallerani18a}. The presence of extra-planar flows may also be the dominant mode for forming the highly supersonic massive MCCs observed in gas-rich star-forming galaxies at $z$\ssim2. For instance, \citet{Swinbank11a} report ISM pressure of $P/k_B\sim$10$^7$\,K\,\cc and 
molecular gas mass of $M_{\rm gas}\simeq10^{8-9}$\,\Msun in the star-forming regions of a gas-rich starburst galaxy at \z\eq2.3 based 
on spatially resolved CO line \obs.

@@

%%If constant column density (Larson's third law) is truly a fundamental properties of molecular clouds,

@@

Properties of the star-forming ISM of galaxies near the Epoch of Reionization are now within reach with ALMA.
While it is possible to obtain sensitive and high fidelity imaging that reveals their gas and dust morphology
on sub-kpc scales and even smaller, such \obs remain time-consuming; logistically, data from multiple cycles pushing to increasing resolution
and sensitivity are needed. As such, observational studies of MCCs on such scales are still missing in the literature.
In this work, we aim to understand the origin and dynamical properties of MCCs in prototypical galaxies at the EoR 
in numerical simulations
to provide a framework within which upcoming \obs can be compared against to aid in the interpretation.

We study the dynamics of MCCs in a \z$\sim$\,6 prototypical galaxy, \flower,
and their temporal evolution using the state-of-the-art cosmological zoom-in simulations (\ncode{Serra}),
which include a chemical network to determine the formation of molecular hydrogen, heating and cooling of the ISM by
UV radiation and metal lines, and detailed stellar feedback.
Properties of \flower are briefly summarized in \Sec{sim}.
We use a three-dimensional clump-finding algorithm to identify MCCs.
We decompose the molecular structures into non-overlapping objects by identifying a set of
density contours with varying densities at multiple evolutionary stages.
Most notably, the accreting phase, starburst phase, and quiescent phase.
We extract properties such as mass, size, Mach number, velocity
dispersion, gas surface density, and virial parameter ($M_{\rm gas}, R, \mathcal{M},
\sigma_{\rm gas}, \Sigma_{\rm gas}, \alpha_{\rm vir}$) for each MCC and perform a Toomre-$Q$ stability analysis on \flower.

The sizes of MCCs are dependent on the choice of $n_{\rm cut}$,
and thus, some of the MCCs within the main disk break into multiple MCCs at higher $n_{\rm cut}$.
The $\sim$200\,pc-scale MCCs found with a low density threshold
correspond to the arms of the disk of \flower and break down into smaller $\lesssim$100\,pc-scale MCCs at higher density thresholds.
Excluding the biggest MCC, the typical mass of MCCs is $10^{6.5}$\,\Msun. 
Similarly massive molecular structures 
have been observed in nearby star-forming and starburst galaxies \citep[e.g.,][]{Keto05a, 
DonovanMeyer13a, Colombo14a, Leroy15a},
and reported in idealized closed-box isolated galaxy simulations done at higher resolution 
(e.g., $l_{\rm cell}\simeq3$\,pc in the 48\,kpc box studied by \citealt{Behrendt16a}); 
however, the IGM and merger and accretion histories are not properly modeled in such simulations.
That said, the similar mass range found in MCCs of \flower\ is
reassuring --- our results are not far off despite the resolution limit ($l_{\rm cell}\simeq$\,30\,pc).
On the other hand, the gas velocity dispersion $\sigma_{\rm gas}$ is rather insensitive to the adopted density threshold.

@@

The variation in $\sigma_{\rm gas}^2/R$ with $\Sigma$ seen in this work is in line with previous studies, suggesting that ``Larson's third law''  
is an artifact arising from observational bias artifact of the limited range of 
column densities a specific molecular line tracer is sensitive to \citep[][]{Ballesteros02a, Ballesteros11a}.
This would imply column density is not a fundamental properties of molecular clouds and not all molecular clouds are virialized.

@@

We perform virial analysis, as motivated by \obs, to assess the stability of MCCs.
% where the virial parameter is used as a surrogate for comparing the potential to kinetic energies of molecule gas structures. 
On average, the virial parameter of MCCs in the starburst phase of \flower is lower than in the accreting phase, as expected for \SF.
Similarly, the virial parameter of MCCs identified at the highest density thresholds are lower than those identified at lower 
density thresholds. 
% Such MCCs are thus susceptible to collapse. 
% The high virial parameter found for some MCCs are found in regions of $Q_{\rm eff}>1$.
% On the other hand, 
Such MCCs identified in (the denser) regions with a lower Toomre $Q_{\rm eff}$ parameter are unstable against collapse.
Star formation is expected to take place within its star-forming {\it clumps} and {\it cores} on smaller scales,
where energy quickly dissipates. This is consistent with the notion that collapsing structures result from
gravitational instability occurring within globally stable structures, which are
supported by turbulence and rotation on large scale.
%%
%Low-$Q$ sets the regions that fragments 
% and within those regions where alphavir is low enough is where structures to further collapse to form stars.
%
%The MCCs of \flower are more massive and turbulent compared to the Milky Way, with 
% higher gas mass surface densities and velocity dispersions. 
This also implies that \obs on spatially resolution better than $\simeq$40\,pc are needed to 
examine the truly star-forming structures (cores),
and thus, \SF in the first galaxies.
%
%% end
%Determining the multi-phase ISM properties of early galaxies
%is a critical piece to understanding the evolution and
%assembly history of galaxies, since they set the pace
%for chemical reactions and excitation rates for the coolants in the ISM (and subsequent star formation).
%While observations leveraging the combination of spatial-spectral imaging of
%multi-band continuum and spectral line emission are crucial,
%% we show in this work that prototypical galaxies at the EoR with properties similar to \flower
%% are expected to have molecular gas structures on sub-kpc scales. 
% Higher resolution imaging of molecular gas 
Such resolution is within reach with ALMA. % 0.01" at the high freq bands
The Next Generation VLA (ngVLA) in the 2030s will enable deeper imaging of the cold gas down to $\sim$\,200\,pc 
via \aco line emission. % 0.005" highest freq; ~0.05 for 16 GHz; CO10 for z~6
On the other hand, cosmological zoom-in simulations, such as \ncode{Serra}, while inherently limited in galaxy
statistics and are subject to the sub-grid models adopted, serve as a useful tool for examining
and making predictions on the morphology and dynamics of the molecular ISM of the first galaxies.
%These predictions can in turn be
%tested with high resolution imaging of the gas content in the first galaxies with telescopes such as ALMA and the ngVLA, thereby
%enabling us to test our findings and refine future simulations to shed light on the physical processes behind
%\SF since the cosmic dark ages.

>>>>

The most massive MCC is excluded in the discussion of MCC dynamics in the remainder of this paper
since its dynamics is dominated by large-scale shear.
Considering only the MCCs identified at the highest density threshold, the mass of MCCs is $\simeq10^{6.5}$\,\Msun.
We note that similarly massive molecular structures (few times 10$^7$\,\Msun\footnote{See \S{3.1} of \citet{Behrendt16a}.}) 
have been reported in idealized closed-box isolated galaxy simulations done at higher resolution (e.g., a maximum resolution of 3\,pc in the 48\,kpc box studied by \citealt{Behrendt16a}). However, the stellar feedback, IGM, merger and accretion histories are not properly modeled in such simulations. That said, the similar mass range found in MCCs of \flower\ is reassuring --- our results are not far off in spite of the limited resolution ($l_{\rm cell}\simeq$\,30\,pc).

@@

If constant column density (Larson's third relation) is truly a fundamental properties of molecular clouds % or the roughly linear relationship between $\rho$ and $R$ of MCCs.
(i.e., if virialization is a universal property of molecular clouds with negligible surface pressure), 
one would expect a single point in the $\sigma_{\rm gas}^2/R$--$\Sigma$ relation at a given gas mass surface density $\Sigma$. 
The variation in $\sigma_{\rm gas}^2/R$ with $\Sigma$ seen here suggests that column density is not a fundamental properties of 
molecular clouds. Previous studies suggest that the observed mass-size relation of molecular clouds 
is solely a result of $\rho\propto M R^{-3}$ and may be an artifact of the limited range of 
column densities a specific molecular line tracer is sensitive to \citep[][]{Ballesteros02a, Ballesteros11a}.
%For instance, \citet{Ballesteros02a} show that the density
%of clouds ($\rho$) is independent of their size ($R$) when one properly extract these quantities in three-dimensional space instead of
%two-dimensional space; the latter is done in \obs.
%When they project the clouds into two-dimensional space and measure the densities and sizes of clouds for a given
%line tracer, they find $\rho \propto R^{-1}$, which is unsurprising since $N_H$ is just a constant.
%They therefore argue that the mass-size relation is simply a result of $\rho\propto M R^{-3}$
%observed in 2D-space at a fixed column density (cf. an actual physical relation of $M \propto R^{4/3}$ claimed in
%observational studies). % where $\rho$ would be proportional to $R^{-1/2}$
%% since $M$ would then scale as $R^{3}$ (ånd naturally display a linear relationship in $\log M - \log R$ plot).

@@

The pressure of MCCs identified at the highest $n_{\rm cut}$ ranges between $\log(P/k_B \,{\rm K}\,{\rm cm}^{-3})\in[7, 9]$, with a median of 7.6. Such high pressure is comparable to those observed in local ULIRGs; however, the molecular clouds in the latter are concentrated within their central regions \citep{Downes98a, Sakamoto08a}. In our simulated galaxy, the high pressure MCCs of \flower are found throughout the disk. This difference likely stems from the different physical mechanisms giving rise to the formation (and thus the nature) of these molecular structures.
%
In local ULIRGs, MCCs are likely formed by shock compression and cloud-cloud collisions that funnel large amounts of gas from the progenitor galaxies towards the central region \citep{Tan00a, Wu18a}. On the other hand, the highly turbulent structures of \flower are the result of extra-planar flows \citep{Kohandel19a} and higher velocity accretion/SN-driven outflows \citep{Gallerani18a}. The presence of extra-planar flows may also be the dominant mode for forming the highly supersonic massive MCCs observed in gas-rich star-forming galaxies at $z$\ssim2. For instance, \citet{Swinbank11a} report ISM pressure of $P/k_B\sim$10$^7$\,K\,\cc and 
molecular gas mass of $M_{\rm gas}\simeq10^{8-9}$\,\Msun in the star-forming regions of a gas-rich starburst galaxy at \z\eq2.3 based 
on spatially resolved CO line \obs.

@@

%%If constant column density (Larson's third relation) is truly a fundamental properties of molecular clouds,

@@

Properties of the star-forming ISM of galaxies near the Epoch of Reionization are now within reach with ALMA. While it is possible to obtain sensitive and high fidelity imaging that reveals their gas and dust morphology on sub-kpc scales and even smaller, such \obs remain time-consuming; logistically, data from multiple cycles pushing to increasing resolution and sensitivity are needed. As such, observational studies of MCCs on such scales are still missing in the literature.
%
In this work, we aim to understand the origin and dynamical properties of MCCs in prototypical galaxies at the EoR in numerical simulations to provide a framework within which upcoming \obs can be compared against to aid in the interpretation.

We study the dynamics of MCCs in a \z$\sim$\,6 prototypical galaxy, \flower, and their temporal evolution using the state-of-the-art cosmological zoom-in simulations (\ncode{Serra}), which include a chemical network to determine the formation of molecular hydrogen, heating and cooling of the ISM by UV radiation and metal lines, and detailed stellar feedback.
%
We use a three-dimensional clump-finding algorithm to identify MCCs. We decompose the molecular structures into non-overlapping objects by identifying a set of H$_2$ density contours with varying densities ($n_{\rm cut}$) at multiple evolutionary stages. Most notably, the accreting phase, starburst phase, and quiescent phase. We extract properties such as mass, size, Mach number, velocity dispersion, gas surface density, and virial parameter ($M_{\rm gas}, R, \mathcal{M}, \sigma_{\rm gas}, \Sigma_{\rm gas}, \alpha_{\rm vir}$) for each MCC and perform a Toomre-$Q$ stability analysis on \flower.

The sizes of MCCs are dependent on the choice of $n_{\rm cut}$, and thus, some of the MCCs within the main disk break into multiple MCCs at higher $n_{\rm cut}$. The $\sim$200\,pc-scale MCCs found with a low density threshold correspond to the arms of the disk of \flower and break down into smaller $\lesssim$100\,pc-scale MCCs at higher density thresholds.
%
Excluding the biggest MCC, the typical mass of MCCs is $10^{6.5}$\,\Msun. Similarly massive molecular structures have been observed in nearby star-forming and starburst galaxies \citep[e.g.,][]{Keto05a, DonovanMeyer13a, Colombo14a, Leroy15a}, and reported in idealized (no stellar feedback) isolated (no cosmological initial conditions) galaxy simulations done at higher resolution (e.g., $l_{\rm cell}\simeq3$\,pc in the 48\,kpc box studied by \citealt{Behrendt16a}). That said, the similar mass range found in MCCs of \flower\ is reassuring --- our results are not far off despite the resolution limit ($l_{\rm cell}\simeq$\,30\,pc). On the other hand, the gas velocity dispersion $\sigma_{\rm gas}$ is rather insensitive to the adopted density threshold.

@@

The variation in $\sigma_{\rm gas}^2/R$ with $\Sigma$ seen in this work is in line with previous studies, suggesting that ``Larson's third relation'' is an artifact arising from observational bias artifact of the limited range of  column densities a specific molecular line tracer is sensitive to \citep[][]{Ballesteros02a, Ballesteros11a}. This would imply column density is not a fundamental properties of molecular clouds and not all molecular clouds are virialized.

@@

% where the virial parameter is used as a surrogate for comparing the potential to kinetic energies of molecule gas structures. 
We perform virial analysis, as motivated by \obs, to assess the stability of MCCs. On average, the virial parameter of MCCs in the starburst phase of \flower is lower than in the accreting phase, as expected for \SF. Similarly, the virial parameter of MCCs identified at the highest density thresholds are lower than those identified at lower 
density thresholds. 
% Such MCCs are thus susceptible to collapse. 
% The high virial parameter found for some MCCs are found in regions of $Q_{\rm eff}>1$.
% On the other hand, 
Such MCCs identified in (the denser) regions with a lower Toomre $Q_{\rm eff}$ parameter are unstable against collapse. Star formation is expected to take place within its star-forming {\it clumps} and {\it cores} on smaller scales, where energy quickly dissipates. This is consistent with the notion that collapsing structures result from gravitational instability occurring within globally stable structures, which are supported by turbulence and rotation on large scale.
%%
%Low-$Q$ sets the regions that fragments 
% and within those regions where alphavir is low enough is where structures to further collapse to form stars.
%
%The MCCs of \flower are more massive and turbulent compared to the Milky Way, with 
% higher gas mass surface densities and velocity dispersions. 
This also implies that \obs on spatially resolution better than $\simeq$40\,pc are needed to examine the truly star-forming structures (cores), and thus, \SF in the first galaxies.
%
%% end
%Determining the multi-phase ISM properties of early galaxies
%is a critical piece to understanding the evolution and
%assembly history of galaxies, since they set the pace
%for chemical reactions and excitation rates for the coolants in the ISM (and subsequent star formation).
%While observations leveraging the combination of spatial-spectral imaging of
%multi-band continuum and spectral line emission are crucial,
%% we show in this work that prototypical galaxies at the EoR with properties similar to \flower
%% are expected to have molecular gas structures on sub-kpc scales. 
% Higher resolution imaging of molecular gas 
Such resolution is within reach with ALMA. % 0.01" at the high freq bands
The Next Generation VLA (ngVLA) in the 2030s will enable deeper imaging of the cold gas down to $\sim$\,200\,pc via \aco line emission. % 0.005" highest freq; ~0.05 for 16 GHz; CO10 for z~6
On the other hand, cosmological zoom-in simulations, such as \ncode{Serra}, while inherently limited in galaxy statistics and are subject to the sub-grid models adopted, serve as a useful tool for examining and making predictions on the morphology and dynamics of the molecular ISM of the first galaxies.
%These predictions can in turn be
%tested with high resolution imaging of the gas content in the first galaxies with telescopes such as ALMA and the ngVLA, thereby
%enabling us to test our findings and refine future simulations to shed light on the physical processes behind
%\SF since the cosmic dark ages.
